SMT-Based Generation of Symbolic Automata \thanks{This work was partially funded by the Associated Team FM4CPS between INRIA and ECNU, Shanghai}

### 2019-03-31 2017-Xudong-SMT/ActaInformatica.tex
Acta Inf. paper: Revised sections 5.2 and 5.3 (Failure Monitor architecture). Significant changes of the characteristic properties of the architecture.

@@

%% CTL
\newcommand{\AG}[1][\ ]{\ensuremath{\mathtt{AG}#1}}
\newcommand{\AF}[1][\ ]{\ensuremath{\mathtt{AF}#1}}
\newcommand{\EF}[1][\ ]{\ensuremath{\mathtt{EF}#1}}
\newcommand{\EX}[1][\ ]{\ensuremath{\mathtt{EX}#1}}
\newcommand{\AW}[3][\ ]{\ensuremath{\mathtt{A}\left[#2\ \mathtt{W}\ #3\right]#1}}
\newcommand{\EU}[3][\ ]{\ensuremath{\mathtt{E}\left[#2\ \mathtt{U}\ #3\right]#1}}

@@

Intuitively, application of the Failure Monitor architecture ensures
that, whenever a failure is registered in the operand component, the
system will be reset, unless a resumption is registered within
$\mathrm{Max}$ time units.  Formally, this statement is decomposed
into several properties, among which 
%
1)~the safety property \emph{``There is always a possibility to reset
  the system after a single failure''} (\ie without additional
failures having to occur in the meantime) specified using CTL as
%
\begin{equation}
  \label{eq:FM:safety}
  \AG
  \bigl(\PortFail \rightarrow
  \EX \EU[]{\lnot \PortFail}{\PortReset}\bigr)
\end{equation}
%
and 2)~the liveness property \emph{``a reset will be fired when asked
  for by a timeout''}:
%
\begin{equation}
  \label{eq:FM:liveness}
  \mathtt{AG}(\PortAsk \rightarrow \mathtt{AF}\ \PortReset).
\end{equation}
%
Additional properties are provided in~\cite{Avocs-RR}.

@@

\begin{figure}[ht]
  \centering
  \includegraphics[width=0.9\columnwidth]{ActaXFIG/Double-OA}
  \caption{The resulting automaton}
  \label{schema:ArchFailure:Double:OA}
\end{figure}

%% In Figure~\ref{schema:ArchFailure:BIP:Double} we show the combination
%% of two Failure Monitor architectures, with their respective controlled
%% processes.
In Figure~\ref{schema:ArchFailure:Double:OA} we show the open
automaton computed for the combined architecture of
Figure~\ref{schema:ArchFailure:BIP:Double}.  It indicates that the two
timers are forced to be synchronous: each can only tick after the
other has also started, and then they will tick, reset, and timeout,
synchronously.
%
\hl{This is a very strong assumption, and we are working on a more
flexible specification (at BIP level) to release this hypothesis while
keeping the safety properties of the composition.}

\noteInSB{The idea was that, by analysing the combination, we observe
  that the architecture does not compose well with itself, generating
  a deadlock (or an implicit deadlock, where only $false$ transitions
  canb be fired.  Based on this observation, we were to improve it
  accodingly.  The problem is that I do not see any deadlocks in
  Figure~\ref{schema:ArchFailure:Double:OA}.  I presume that some of
  the asterisks should be only $false$ or something of a kind.}

>>>>

\usepackage{tikz}
\usetikzlibrary{matrix,arrows,decorations.pathmorphing,shapes}
\usetikzlibrary{decorations.markings}
\usetikzlibrary{calc,positioning}

%% \tikzset{
%%   negated/.style={
%%     decoration={
%%       markings,
%%       mark=at position 0.5 with {\node[transform shape] (tempnode) {$/$};},
%%     },
%%     postaction={decorate},
%%   },
%% }

%% \tikzset{label/.style={text=blue,font=\normalsize}}
%% \tikzset{box/.style={fill=yellow,draw=brown}}
%% \tikzset{port north/.style={fill=white,draw=brown,anchor=north,minimum size=1.2em}}
%% \tikzset{port east/.style={fill=white,draw=brown,anchor=east}}
%% \tikzset{port west/.style={fill=white,draw=brown,anchor=west}}
%% \tikzset{port south/.style={fill=white,draw=brown,anchor=south}}

\newenvironment{tikzlts}[1][15mm]{\begin{tikzpicture}
[->,>=stealth',shorten >=1pt,auto,node distance={#1},thick,draw=black,
state/.style={ellipse,draw=black,font=\sffamily},align=center,text=black]} %,fill=blue!20 \large\bfseries
{\end{tikzpicture}}

@@

%% CTL
\newcommand{\AG}[1][\ ]{\ensuremath{\mathtt{AG}#1}}
\newcommand{\AF}[1][\ ]{\ensuremath{\mathtt{AF}#1}}
\newcommand{\EF}[1][\ ]{\ensuremath{\mathtt{EF}#1}}
\newcommand{\EX}[1][\ ]{\ensuremath{\mathtt{EX}#1}}
\newcommand{\AW}[3][\ ]{\ensuremath{\mathtt{A}\left[#2\ \mathtt{W}\ #3\right]#1}}
\newcommand{\EU}[3][\ ]{\ensuremath{\mathtt{E}\left[#2\ \mathtt{U}\ #3\right]#1}}
\newcommand{\AU}[3][\ ]{\ensuremath{\mathtt{A}\left[#2\ \mathtt{U}\ #3\right]#1}}

@@

Notice that the model in Figure~\ref{schema:ArchFailure:BIP} allows
$\PortReset$ to happen in any state.  However, it is implied that,
when the coordinator $\NameCtrl$ is in state $\cstate{2}$, this
$\PortReset$ is ``asked for'' by the architecture (it follows the
firing of $\PortAsk$), whereas in other states the corresponding
transition reflects an ``external'' reset.  Although this can be
formalised in the model using data variables, we have chosen not to do
so for the sake of simplicity.

Intuitively, application of the Failure Monitor architecture ensures
that, whenever a failure is registered in the operand component, the
system will be reset, unless a resumption is registered within
$\mathrm{Max}$ time units.  Formally, this statement is decomposed
into several properties, among which 
%
1)~the safety property \emph{``A reset is never asked for unless a failure occurs''} specified using CTL as
%
\begin{equation}
  \label{eq:FM:safety}
  %% \AG
  %% \bigl(\PortFail \rightarrow
  %% \EX \EU[]{\lnot \PortFail}{\PortReset}\bigr)
  \AW[]{\lnot \PortAsk}{\PortFail}
\end{equation}
%
and 2)~the liveness property \emph{``in the absence of an external
  reset, a reset can always be asked for after a non-transient
  failure''}:
%
\begin{equation}
  \label{eq:FM:liveness}
  %% \mathtt{AG}(\PortAsk \rightarrow \mathtt{AF}\ \PortReset).
  \AG \bigl(
  \PortFail \rightarrow \EU[]{(\lnot \PortReset \land \lnot \PortResume)}{\PortAsk}
  \bigr)
\end{equation}
%
Additional properties are provided in~\cite{Avocs-RR}.

@@

\begin{figure}[ht]
  \centering
  %% \includegraphics[width=0.9\columnwidth]{ActaXFIG/Double-OA}
  \resizebox{\columnwidth}{!}{
  \begin{tikzlts}
    \node[state] (0000) {0000};
    \node[state, above right= 10mm and 30mm of 0000] (1010) {1010};
    \node[state, below right= 10mm and 30mm of 0000] (0101) {0101};
    \node[state, below right= 10mm and 30mm of 1010] (1111) {1111};
    \node[state, above right= 15mm and 20mm of 1010] (2200) {2200};

    \path[draw,->]
    (0000.135)
    .. controls +(-7mm,3mm) and +(-3mm,7mm) ..
    node {reset}
    (0000.120)
    ;

    \path
    (0000)
    edge[loop left] node[left] {resume1(*)} (0000)
    edge[bend left] node[above] {fail1(true)} (1010)
    edge[bend right] node[below] {fail2(true)} (0101)
    ;

    \path[draw,->]
    (0000.-120)
    .. controls +(-3mm,-7mm) and +(-7mm,-3mm) ..
    node[below left] {resume2(*)}
    (0000.-135)
    ;

    \path[draw,->]
    (1111.60)
    .. controls +(3mm,7mm) and +(7mm,3mm) ..
    node[above right] {fail1(*)}
    (1111.45)
    ;

    \path
    (1111)
    edge node {reset} (0000)
    edge[loop right] node[right] {fail2(*)} (1111)
    edge[out=170,in=-20] node[left] {resume2(true)} (1010)
    edge[out=-170,in=20] node[left] {resume1(true)} (0101)
    ;

    \path[draw,->]
    (1111.-45)
    .. controls +(7mm,-3mm) and +(3mm,-7mm) ..
    node[below right] {\begin{tabular}{@{}l@{}}
        tick\\
        {}[t1<Max1 \& t2<Max2]\\
        \{t1:=t1+1; t2:=t2+1\}
    \end{tabular}}
    (1111.-60)
    ;

    \path[draw,->]
    (1111) edge
    node[above right] {\begin{tabular}{@{}l@{}}
        ask\\
        {}[t1>=Max1 \& t2>=Max2]
    \end{tabular}}
    (2200)
    ;

    \path
    (1010)
    edge[out=-170,in=30] node[above] {reset} (0000)
    edge[out=-150,in=20] node[below] {resume1(true)} (0000)
    edge[bend left] node[above] {fail2(true)} (1111)
    ;

    \path[draw,->]
    (1010.45)
    .. controls +(7mm,3mm) and +(3mm,7mm) ..
    node[above right] {fail1(*)}
    (1010.80)
    ;

    \path[draw,->]
    (1010.100)
    .. controls +(-3mm,7mm) and +(-7mm,3mm) ..
    node[above left] {resume2(*)}
    (1010.135)
    ;

    \path
    (0101)
    edge[out=170,in=-30] node[below] {reset} (0000)
    edge[out=150,in=-20] node[above] {resume2(true)} (0000)
    edge[bend right] node[below right] {fail1(true)} (1111)
    ;

    \path[draw,->]
    (0101.-45)
    .. controls +(7mm,-3mm) and +(3mm,-7mm) ..
    node[below right] {fail2(*)}
    (0101.-80)
    ;

    \path[draw,->]
    (0101.-100)
    .. controls +(-3mm,-7mm) and +(-7mm,-3mm) ..
    node[below left] {resume1(*)}
    (0101.-135)
    ;

    \path[draw,->]
    (2200.west)
    .. controls +(left:5cm) ..
    node[above left] {reset}
    (0000.80)
    ;

    \path[draw,->]
    (2200.0)
    .. controls +(7mm,0) and +(3mm,-7mm) ..
    node[right] {fail2(*)}
    (2200.-20)
    ;

    \path[draw,->]
    (2200.40)
    .. controls +(3mm,7mm) and +(7mm,3mm) ..
    node[above right] {fail1(*)}
    (2200.20)
    ;

    \path[draw,->]
    (2200.100)
    .. controls +(-3mm,7mm) and +(3mm,7mm) ..
    node[above] {resume2(*)}
    (2200.80)
    ;

    \path[draw,->]
    (2200.160)
    .. controls +(-7mm,3mm) and +(-3mm,7mm) ..
    node[left] {resume1(*)}
    (2200.150)
    ;
  \end{tikzlts}}
  \caption{The resulting automaton}
  \label{schema:ArchFailure:Double:OA}
\end{figure}

%% In Figure~\ref{schema:ArchFailure:BIP:Double} we show the combination
%% of two Failure Monitor architectures, with their respective controlled
%% processes.
In Figure~\ref{schema:ArchFailure:Double:OA} we show the open
automaton computed for the combined architecture of
Figure~\ref{schema:ArchFailure:BIP:Double}.
%
Comparing the guards of the transitions $\PortTick$ and $\PortAsk$
leaving state $1111$, we see that $\PortAsk$ can never be taken,
unless $\mathrm{Max1} = \mathrm{Max2}$.  This illustrates the
\emph{interference} phenomenon discussed in
\cite{AttieBBJS16-architectures-faoc}, showing that the liveness
property (\ref{eq:FM:liveness}) is not preserved when two copies of
the Failure Monitor architecture are composed together.\footnote{%
%
  One of the main results of \cite{AttieBBJS16-architectures-faoc}
  states that in a system obtained by an application of several
  architectures, liveness properties are preserved if these
  architectures are \emph{pair-wise non-interfering}.  However, no
  results were provided to check whether two are architectures are
  interfering or not.
%
} In this example, the interference is caused by the strong
synchronisation of the $\PortTick$ ports of the two architectures.  It
can be avoided by replacing the \mbox{$\PortTick
  \bullet\!\!\!-\!\!\!-\!\!\!\bullet \NameTimer.\PortTick$} connector
in Figure~\ref{schema:ArchFailure:BIP} by \mbox{$\PortTick
  \blacktriangleright\!\!\!-\!\!\!-\!\!\!\bullet
  \NameTimer.\PortTick$} and applying the so-called \emph{maximal
  progress} whereby, whenever possible, the interaction $\{\PortTick,
\NameTimer.\PortTick\}$ is preferred to the firing of $\PortTick$
alone.  However, this fix goes beyond the scope of the current paper.

%% It indicates that the two
%% timers are forced to be synchronous: each can only tick after the
%% other has also started, and then they will tick, reset, and timeout,
%% synchronously.
%% %
%% \hl{This is a very strong assumption, and we are working on a more
%% flexible specification (at BIP level) to release this hypothesis while
%% keeping the safety properties of the composition.}

%% \noteInSB{The idea was that, by analysing the combination, we observe
%%   that the architecture does not compose well with itself, generating
%%   a deadlock (or an implicit deadlock, where only $false$ transitions
%%   canb be fired.  Based on this observation, we were to improve it
%%   accodingly.  The problem is that I do not see any deadlocks in
%%   Figure~\ref{schema:ArchFailure:Double:OA}.  I presume that some of
%%   the asterisks should be only $false$ or something of a kind.}
